---
title: 'Peer Graded Assigment: Course Project 1'
author: "Héctor Morales Osorio"
date: "Sunday, September 25, 2016"
output: md_document
keep: md_document
--- 
Dataset: Activity moninotoring data

Variables:
·steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)

·date: The date on which the measurement was taken in YYYY-MM-DD format

·interval: Identifier for the 5-minute interval in which measurement was taken

## 1. Code for reading in the dataset and/or processing the data

```{r}
setwd("D:/")
data<-read.csv("activity.csv" )
head(data)
```

Structure data
```{r}
str(data)
summary(data)
```

Transform variable date
```{r}
data$date<-as.Date(as.character(data$date),format="%Y-%m-%d")
```

## 2. Histogram of the total number of steps taken each day
```{r}
stepsByDay<-data.frame(sapply(split(data$steps,data$date),sum))
hist(stepsByDay[,1],xlab="Number of step",main="Histogram of the total number of steps taken each day")
```

## 3. Mean and median number of steps taken each day
```{r}
meanStepsByDay<-data.frame(sapply(split(data$steps,data$date),mean))
colnames(meanStepsByDay)<-"meanStepsByDay"
meanStepsByDay
medianStepsByDay<-data.frame(sapply(split(data$steps,data$date),median))
colnames(medianStepsByDay)<-"medianStepsByDay"
medianStepsByDay

```

## 4. Time series plot of the average number of steps taken

```{r}
meanStepsInterval<-data.frame(sapply(split(data$steps,data$interval),mean))
colnames(meanStepsInterval)[1]<-"meanStepsInterval"

plot(rownames(meanStepsInterval),meanStepsInterval[,1],type="l",xlab="Interval",ylab="Number of steps",main="Average number of steps taken",ylim=c(0,200))
```

## 5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
maxIntervalAverage<-max(na.omit(meanStepsInterval[,1]))
subset(meanStepsInterval,meanStepsInterval==maxIntervalAverage)

```

## 6. Code to describe and show a strategy for imputing missing data
Number of missing values in the dataset 

```{r}
NAdata<-na.omit(data)
dim(data)[1]-dim(NAdata)[1]
```
Replace NA for the mean 
```{r}
meanIntervalAverage<-mean(na.omit(meanStepsByDay[,1]))
data[is.na(data)]<-meanIntervalAverage
```
## 7. Histogram of the total number of steps taken each day 

```{r}
stepsByDay2<-data.frame(sapply(split(data$steps,data$date),sum))
hist(stepsByDay2[,1],xlab="Number of steps", main="Histogram of the total number of steps taken each day")

```

The mean and median total number of steps taken per day. 
```{r}
meanStepsByDay2<-data.frame(sapply(split(data$steps,data$date),mean))
colnames(meanStepsByDay2)[1]<-"meanStepsByDay"
meanStepsByDay2
medianStepsByDay2<-data.frame(sapply(split(data$steps,data$date),median))
colnames(medianStepsByDay2)[1]<-"medianStepsByDay"
medianStepsByDay2

```

## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
par(mfrow = c(2, 1))
days<-weekdays(data$date, abbreviate = FALSE)
weekend<-replace(days,days=="Saturday"|days=="Sunday","Weekend")
weekday<-replace(weekend,weekend=="Monday"|weekend=="Tuesday"|weekend=="Wednesday  
                        "| weekend=="Thursday"|weekend=="Friday","Weekday")
data2<-cbind(data,weekday)
weekend2<-subset(data2,weekday=="Weekend")
meanWeekend<-data.frame(sapply(split(weekend2$steps,weekend2$interval),mean))
plot(rownames(meanWeekend),meanWeekend[,1],type="l",xlab="Interval",ylab="Number of steps",main="Weekend",ylim=c(0,200))

weekday2<-subset(data2,weekday=="Weekday")
meanWeekday<-data.frame(sapply(split(weekday2$steps,weekday2$interval),mean))
plot(rownames(meanWeekday),meanWeekday[,1],type="l",xlab="Interval",ylab="Number of steps",main="Weeday",ylim=c(0,200))

```

